---
title: "Housing Cost Analysis"
format: html
editor: visual
---

```{r}
house_data <- read.csv("C:/Users/Carlos/Downloads/kc_house_data.csv")
head(house_data)

```

Linear Model that uses the predictors to predict price.

Data Cleaning: Lets properly format the date column in to YY/MM/DD and rename it to date_purchased

```{r}
library(dplyr)
#house_data$parsed_date <- as.Date(house_data$date, format = "%Y%m%dT%H%M%S")
house_data$date <- NULL
house_data <- house_data %>% rename(date_purchased = parsed_date)

```

Creating our linear model. We will use step-wise regression to remove unecessarily predictors, Ocam's Razor style.

```{r}
# Latitude and Longitute have been removed because they are dependent on eachother.
# date_purchased has been removed since it is actually a categorical data type. 

housefull_lm <- lm(price ~ sqft_living + sqft_lot + bedrooms + bathrooms + grade + floors + waterfront + view + condition + sqft_above + sqft_basement + yr_built + yr_renovated + sqft_living15 + sqft_lot15,
   data = house_data,
   na.action = na.omit)

#summary(housefull_lm)

# We will perform a stepwise regression to remove variables that don't have much of an impact on the final regression. 
library(MASS)
step <- stepAIC(housefull_lm, direction = "both")
step
```

The stepwise regression removed the predictors Sqft_Lot, SqFt_Basement, and SqFt_Above.\
This is likely because these variables are represented by the more influential sqft_living, sqft_living15, and sqft_lot15 predictors.

We will incorporate a weight to records that were recently sold. Recently sold homes are reflective of the current trends more than houses sold a long time ago. We will use date_purchased to determine the weights of each record.

```{r}
library(lubridate)
house_data$Year = year(house_data$date_purchased)
house_data$Weight = house_data$Year - 2016

```
